---
title: "Pladas_R_Package"
author: "Shawn Pladas"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation

Load in common libraries used by this package

```{r cars}
library(ggplot2)
library(tidyverse)
library(dplyr)
```

Install Pladas_R_Package 

```{r eval=FALSE}
devtools::install_github("shawn-pladas/R_package_Pladas")
library(PladasRPackage)
```

Download some data for the package:

```{r}
download.file(url = "https://github.com/shawn-pladas/R_package_Pladas/blob/master/data/Bat_Capture_Data.csv", destfile = "../data/bats.csv")
```

##Working Example

+An Example of each function and how it works
+Say something about why this is a good thing to do
+say something about expected output

na.remove removes NA values from your entire data frame across all rows and columns. Useful for a quick and immediate way to eliminate NA values in data. Provides a data frame free of NA values. See example below:

```{r}

bats1 <- na.remove(bats)

```

calc_diff takes two columns with numeric values and calculates the difference between them. Useful for if you need to add a column of calculated difference in weights, size, etc. Provides a column within a dataframe of a numeric difference. 

```{r}

bats2 <- calc_diff(bats1, mass_bat_bag, mass_bag)

```

The histo_plot function provides a bar graph of the x and y input data and includes a parameter to add in a title. Useful for quick representation of values within a data set. 

```{r}

histo_plot(bats2, sex, location, "Sex ratio by location")

```

lm_mult_predictor creates linear model comparing two predictor columns, numerical and categorical, to response column. Useful for quick representation and visualization of linear models with multiple predictors. Parameters include a data frame, xcol for a numerical predictor, ycol for a response variable, and zcol for a categorical predictor. Also, provides a parameter for including a title. 

```{r}

mult_lm_predictor(bats2, difference, forearmlength_mm, sex, "Weight and sex as predictors of forearm length")

```

Lastly, ANCOVA create a multiple predictor linear model analysis using numerical and categorical data. Parameters include a data frame, "response" a response variable, "predictor1" a numerical predictor, "predictor2" a categorical predictor variable. Also, provides a parameter for including a title. Returns linear models tests including T-test, R^2 values, p-values, etc. 

```{r}

ancova(bats3, "difference", "forearmlength_mm", "sex")

```



